``FISTA'' in Banach spaces with adaptive discretisations

### 2021-09-22 paper.tex
A few notes to improve

\let\F\mathds\let\C\mathcal\newcommand{\R}{\F{R}}\newcommand{\A}{\tens{A}}
\newcommand{\norm}[1]{{\left\lVert #1 \right\rVert}}
\newcommand{\Norm}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}
\newcommand{\IP}[2]{\left\langle #1,#2 \right\rangle}
\newcommand{\ip}[2]{#1 \vcenter{\hbox{\resizebox{6pt}{!}{\ensuremath\cdot}}} #2}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\splitln}[4]{\begin{cases} #1 & #2 \\ #3 & #4\end{cases}}
\newcommand{\1}{\F{1}}
\DeclareMathOperator{\st}{\;s.t.\;}
\renewcommand{\epsilon}{\varepsilon}\renewcommand{\bar}{\overline}
\renewcommand{\hat}{\widehat}\renewcommand{\tilde}{\widetilde}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sign}{sign}
\newcommand{\Emin}[1][\varf0]{\inf_{#1\in\F H}\op{E}(#1)}

@@

	Without loss of generality we will assume $L=1$, i.e. $\nabla\op{f}$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\op{E}$ with $\frac{\op{E}}{L}$.\edit{ A more important distinction is to say that the properties of $\op{f}$ and $\op{g}$ are stated with respect to the Hilbert norm.}{} In particular,
	\begin{equation}
		\norm{\nabla\op{f}(\varf0)- \nabla\op{f}(\varf1)} \leq \norm{\varf0-\varf1}
	\end{equation}
	for all $\varf0,\varf1\in\F H$ and $\op{g}$ is called `simple' if \edit{}{it is proper, convex, weakly lower-semicontinuous, and}
	\begin{equation}
		\argmin_{\varf0\in\tilde{\spcf0}}\tfrac12\norm{\varf0-\varf1}^2 + \op{g}(\varf0)
	\end{equation}
	is exactly computable for all $\varf1\in\F H$ and all $\tilde{\spcf0}\in\{\spcf0^n\}_{n=0}^\infty$. \edit{}{We denote the orthogonal projection operators $\tens{\Pi}_n\colon\F H\to\spcf0^n$ such that $\tens{\Pi}_n\varf0-\varf0 \perp \varf1$ for all $\varf0\in\F H$, $\varf1\in\spcf0^n$.}

@@

	\subsubsection{Single iterations}
	We first wish to understand a single iteration of Algorithm~\ref{alg: refining FISTA}. This is done through the following two lemmas.
	\begin{lemma}[equivalent to {\cite[Lemma \edit{1}{3.1}]{Chambolle2015}}]\label{thm: descent lemma}
		\Copy{thm: descent lemma}{
			Suppose $\nabla \op{f}$ is 1-Lipschitz, for any $\bar{\varf0}\in \edit{\spcf0^{n}}{\F H}$ define
			$$ \varf0 \coloneqq \argmin_{\varf0\in \spcf0^{n}} \tfrac12\norm{\varf0-\bar{\varf0}+\nabla \op{f}(\bar{\varf0})}^2+\op{g}(\varf0).$$
			Then, for all $\varf2\in\edit{\bar{(\spcf0^n)^*}\supset\F H}{\spcf0^n}$, we have
			\edit{\\\centerline{$\displaystyle \op{E}(\varf0) + \tfrac12\norm{\varf0-\tens{\Pi}_n\varf2}^2 \leq \op{E}(\tens{\Pi}_n\varf2) + \tfrac12\norm{\bar{\varf0}-\tens{\Pi}_n\varf2}^2.$}}{
			$$\op{E}(\varf0) + \tfrac12\norm{\varf0-\varf2}^2 \leq \op{E}(\varf2) + \tfrac12\norm{\tens{\Pi}_n\bar{\varf0}-\varf2}^2 \leq \op{E}(\varf2) + \tfrac12\norm{\bar{\varf0}-\varf2}^2.$$}
			\edit{where $\tens{\Pi}_n\colon\bar{(\spcf0^n)^*}\to\spcf0^n$ is the orthogonal projection.}{}
		}
	\end{lemma}
	\noindent This is exactly the result of \cite{Chambolle2015} applied to the function $\varf0\mapsto\op{E}(\tens{\Pi}_n\varf0)$. Applying Lemma~\ref{thm: descent lemma} to the iterates from Algorithm~\ref{alg: refining FISTA} gives a more explicit inequality.

	\begin{lemma}[{\edit{analogous to }{}\cite[\edit{Theorem 2}{(17)}]{Chambolle2015}, \cite[\edit{Theorem 1}{Lmma 4.1}]{Beck2009}}]\label{thm: one step FISTA}
		\Copy{thm: one step FISTA}{
			Let \edit{$\spcf0^n\subset\F H\cap\spcf0$, }{}$\varf2_n\in\spcf0^n$ be chosen arbitrarily and $\varf0_n$/$\varf1_n$ be generated by Algorithm~\ref{alg: refining FISTA} for all $n\in\F N$.
			For all $n\in\F N$, it holds that}
		\begin{equation}\Copy{thm:eq: one step FISTA}{
			t_{n}^2(\op{E}(\varf0_{n}) - \op{E}(\varf2_n)) - (t_{n}^2-t_{n})(\op{E}(\varf0_{n-1})-\op{E}(\varf2_n)) \leq \tfrac1{2}\left[\norm{\varf1_{n-1}}^2-\norm{\varf1_{n}}^2\right] + \IP{\varf1_{n}-\varf1_{n-1}}{\varf2_n}.}\label{eq: one-step FISTA}
		\end{equation}
	\end{lemma}
	\noindent The proof is given in Theorem~\ref{app:thm: one step FISTA} and is a result of the convexity of $\op{E}$ for a well chosen $\varf2$ in Lemma~\ref{thm: descent lemma}.

@@

	\begin{theorem}[Lemma~\ref{thm: one step FISTA}]\label{app:thm: one step FISTA}
		\Paste{thm: one step FISTA}
		\begin{equation}
			\Paste{thm:eq: one step FISTA}
			\label{app:eq: one-step FISTA}
		\end{equation}
	\end{theorem}
	\begin{proof}
		Modifying \cite[\edit{Theorem 2}{Thm 3.2}]{Chambolle2015}, for $n\geq1$ we apply Lemma~\ref{thm: descent lemma} with $\bar{\varf0} = \bar{\varf0}_{n-1}$ and $\varf2 = (1-\frac{1}{t_{n}})\varf0_{n-1} + \frac{1}{t_{n}}\varf2_n\edit{}{\in\spcf0^n}$. This gives
		\begin{equation}
			\op{E}(\varf0_{n}) + \tfrac12\norm{\tfrac{1}{t_{n}}\varf1_{n}-\tfrac{1}{t_{n}}\varf2_n}^2 \leq \op{E}\left((1-\tfrac{1}{t_{n}})\varf0_{n-1}+\tfrac{1}{t_{n}}\varf2_n\right) + \tfrac12\norm{\tfrac{1}{t_{n}}\varf1_{n-1}-\tfrac{1}{t_{n}}\varf2_n}^2.
		\end{equation}
		By the convexity of $\op{E}$, this reduces to
		\begin{equation}
			\op{E}(\varf0_{n}) - \op{E}(\varf2_n) - (1-\tfrac{1}{t_{n}})[\op{E}(\varf0_{n-1})-\op{E}(\varf2_n)] \leq \tfrac1{2t_{n}^2}\norm{\varf1_{n-1}-\varf2_n}^2 - \tfrac1{2t_{n}^2}\norm{\varf1_{n}-\varf2_n}^2 = \tfrac1{2t_{n}^2}\left[\norm{\varf1_{n-1}}^2-\norm{\varf1_{n}}^2\right] + \tfrac1{t_{n}^2}\IP{\varf1_{n}-\varf1_{n-1}}{\varf2_n}.
		\end{equation}
		\edit{}{Multiplying through by $t_n^2$ gives the desired inequality.}
		\qed\end{proof}

>>>>

\let\F\mathds\let\C\mathcal\newcommand{\R}{\F{R}}\newcommand{\A}{\tens{A}}
\newcommand{\norm}[1]{{\left\lVert #1 \right\rVert}}
\newcommand{\Norm}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}
\newcommand{\IP}[2]{\left\langle #1,#2 \right\rangle}
\newcommand{\ip}[2]{#1 \vcenter{\hbox{\resizebox{6pt}{!}{\ensuremath\cdot}}} #2}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\splitln}[4]{\begin{cases} #1 & #2 \\ #3 & #4\end{cases}}
\newcommand{\1}{\F{1}}
\DeclareMathOperator{\st}{\;s.t.\;}
\renewcommand{\epsilon}{\varepsilon}\renewcommand{\bar}{\overline}
\renewcommand{\hat}{\widehat}\renewcommand{\tilde}{\widetilde}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sign}{sign}
\newcommand{\Emin}[1][\varf0]{\inf_{#1\in\F H}\op{E}(#1)}
\newcommand{\todo}[1]{{\color{red}#1}}

@@

	Without loss of generality we will assume $L=1$, i.e. $\nabla\op{f}$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\op{E}$ with $\frac{\op{E}}{L}$.\edit{ A more important distinction is to say that the properties of $\op{f}$ and $\op{g}$ are stated with respect to the Hilbert norm.}{} In particular,
	\begin{equation}
		\norm{\nabla\op{f}(\varf0)- \nabla\op{f}(\varf1)} \leq \norm{\varf0-\varf1}
	\end{equation}
	for all $\varf0,\varf1\in\F H$ and $\op{g}$ is called `simple' if \edit{}{it is proper, convex, weakly lower-semicontinuous, and}
	\begin{equation}
		\argmin_{\varf0\in\tilde{\spcf0}}\tfrac12\norm{\varf0-\varf1}^2 + \op{g}(\varf0)
	\end{equation}
	is exactly computable for all $\varf1\in\F H$ and all $\tilde{\spcf0}\in\{\spcf0^n\}_{n=0}^\infty$. \edit{}{We denote the orthogonal projection operators $\tens{\Pi}_n\colon\F H\to\spcf0^n$ such that $\tens{\Pi}_n\varf0-\varf0 \perp \varf1$ for all $\varf0\in\F H$, $\varf1\in\spcf0^n$. \todo{Finite dimensional?}}

@@

	\subsubsection{Single iterations}
	We first wish to understand a single iteration of Algorithm~\ref{alg: refining FISTA}. This is done through the following two lemmas.
	\begin{lemma}[equivalent to {\cite[Lemma \edit{1}{3.1}]{Chambolle2015}}]\label{thm: descent lemma}
		\Copy{thm: descent lemma}{
			Suppose $\nabla \op{f}$ is 1-Lipschitz, for any $\bar{\varf0}\in \edit{\spcf0^{n}}{\F H}$ define
			$$ \varf0 \coloneqq \argmin_{\varf0\in \spcf0^{n}} \tfrac12\norm{\varf0-\bar{\varf0}+\nabla \op{f}(\bar{\varf0})}^2+\op{g}(\varf0).$$
			Then, for all $\varf2\in\edit{\bar{(\spcf0^n)^*}\supset\F H}{\spcf0^n}$, we have
			\edit{\\\centerline{$\displaystyle \op{E}(\varf0) + \tfrac12\norm{\varf0-\tens{\Pi}_n\varf2}^2 \leq \op{E}(\tens{\Pi}_n\varf2) + \tfrac12\norm{\bar{\varf0}-\tens{\Pi}_n\varf2}^2.$}}{
			$$\op{E}(\varf0) + \tfrac12\norm{\varf0-\varf2}^2 \leq \op{E}(\varf2) + \tfrac12\norm{\tens{\Pi}_n\bar{\varf0}-\varf2}^2 \leq \op{E}(\varf2) + \tfrac12\norm{\bar{\varf0}-\varf2}^2.$$}
			\edit{where $\tens{\Pi}_n\colon\bar{(\spcf0^n)^*}\to\spcf0^n$ is the orthogonal projection.}{}
		}
	\end{lemma}
\todo{check this}
	\noindent This is exactly the result of \cite{Chambolle2015} applied to the function $\varf0\mapsto\op{E}(\tens{\Pi}_n\varf0)$. Applying Lemma~\ref{thm: descent lemma} to the iterates from Algorithm~\ref{alg: refining FISTA} gives a more explicit inequality.

	\begin{lemma}[{\edit{analogous to }{}\cite[\edit{Theorem 2}{(17)}]{Chambolle2015}, \cite[\edit{Theorem 1}{Lemma 4.1}]{Beck2009}}]\label{thm: one step FISTA}
		\Copy{thm: one step FISTA}{
			Let \edit{$\spcf0^n\subset\F H\cap\spcf0$, }{}$\varf2_n\in\spcf0^n$ be chosen arbitrarily and $\varf0_n$/$\varf1_n$ be generated by Algorithm~\ref{alg: refining FISTA} for all $n\in\F N$.
			For all $n\in\F N$, it holds that}
		\begin{equation}\Copy{thm:eq: one step FISTA}{
			t_{n}^2(\op{E}(\varf0_{n}) - \op{E}(\varf2_n)) - (t_{n}^2-t_{n})(\op{E}(\varf0_{n-1})-\op{E}(\varf2_n)) \leq \tfrac1{2}\left[\norm{\varf1_{n-1}}^2-\norm{\varf1_{n}}^2\right] + \IP{\varf1_{n}-\varf1_{n-1}}{\varf2_n}.}\label{eq: one-step FISTA}
		\end{equation}
	\end{lemma}
	\noindent The proof is given in Theorem~\ref{app:thm: one step FISTA} and is a result of the convexity of $\op{E}$ for a well chosen $\varf2$ in Lemma~\ref{thm: descent lemma}.

@@

	\begin{theorem}[Lemma~\ref{thm: one step FISTA}]\label{app:thm: one step FISTA}
		\Paste{thm: one step FISTA}
		\begin{equation}
			t_{n}^2(\op{E}(\varf0_{n}) - \op{E}(\varf2_n)) - (t_{n}^2-t_{n})(\op{E}(\varf0_{n-1})-\op{E}(\varf2_n)) \leq \tfrac1{2}\left[\norm{\varf1_{n-1}}^2-\norm{\varf1_{n}}^2\right] + \IP{\varf1_{n}-\varf1_{n-1}}{\varf2_n}.
			\label{app:eq: one-step FISTA}
		\end{equation}
	\end{theorem}
	\begin{proof}
		Modifying \cite[\edit{Theorem 2}{Thm 3.2}]{Chambolle2015}, for $n\geq1$ we apply Lemma~\ref{thm: descent lemma} with $\bar{\varf0} = \bar{\varf0}_{n-1}$ and $\varf2 = (1-\frac{1}{t_{n}})\varf0_{n-1} + \frac{1}{t_{n}}\varf2_n\edit{}{\in\spcf0^n}$. This gives
		\begin{equation}
			\op{E}(\varf0_{n}) + \tfrac12\norm{\tfrac{1}{t_{n}}\varf1_{n}-\tfrac{1}{t_{n}}\varf2_n}^2 \leq \op{E}\left((1-\tfrac{1}{t_{n}})\varf0_{n-1}+\tfrac{1}{t_{n}}\varf2_n\right) + \tfrac12\norm{\tfrac{1}{t_{n}}\varf1_{n-1}-\tfrac{1}{t_{n}}\varf2_n}^2.
		\end{equation}
		By the convexity of $\op{E}$, this reduces to
		\begin{equation}
			\op{E}(\varf0_{n}) - \op{E}(\varf2_n) - (1-\tfrac{1}{t_{n}})[\op{E}(\varf0_{n-1})-\op{E}(\varf2_n)] \leq \tfrac1{2t_{n}^2}\norm{\varf1_{n-1}-\varf2_n}^2 - \tfrac1{2t_{n}^2}\norm{\varf1_{n}-\varf2_n}^2 = \tfrac1{2t_{n}^2}\left[\norm{\varf1_{n-1}}^2-\norm{\varf1_{n}}^2\right] + \tfrac1{t_{n}^2}\IP{\varf1_{n}-\varf1_{n-1}}{\varf2_n}.
		\end{equation}
		\edit{}{Multiplying through by $t_n^2$ gives the desired inequality.}
		\qed\end{proof}
